---
title: "Problem Set 2 - PPOL 6802 Spring 2024"
author: "INSERT NAME"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Set #2

Here is where you will provide the answers to problem set 2. I will provide you with the basic "buckets" to put everything in to make the R Markdown creation process as straightforward as possible. But remember that if you run into any issues, ChatGPT can be extremely helpful!

### Loading the Data

To complete this problem set, you will need to download the necessary data from Github. You can do that by running the code below:

```{r message = FALSE}
library(tidyverse)

d1 <- read_csv("https://raw.githubusercontent.com/alexlundry/ppol6802_spring24/main/problem_sets/datasets/pset1_county_data.csv")

d2 <- read_csv("https://raw.githubusercontent.com/cphalpert/census-regions/master/us%20census%20bureau%20regions%20and%20divisions.csv")

d3 <- left_join(d1, d2, by = "State")
```

### Question 1 - Bar Chart 

Using ggplot2, make a bar chart of the number of counties in each census Division.

**Bonus (not required)**: Sort the bars in order (+1 point)

```{r}
# insert and run ggplot2 code for question 1 here.

```

### Question 2 - Histogram

Using ggplot2, create a histogram of `pct_bachelors` that is split into 20 bins.

```{r}
# insert and run your ggplot code for question 2 here
```

### Question 3 - Bar Chart Reproduction

Using ggplot2, reproduce the bar chart below. To help you do this, here are a few key notes:

-   You must first count the counties (rows) within each `Division`
-   Then the percentage is calculated; it is rounded to 2 decimal places
-   These are horizontal bars.
-   The bars are ordered descending by their percent.
-   The bars are colored by `Region`
-   The bars are labeled with their percent, and the position of that label is on the inside end of the bar
-   I am manually setting colors.  They are "deepskyblue", "burlywood", "brown1", and "lightgreen".
-   The axis labels are formatted to be in %
-   There is a title and labels.

```{r}
# insert and run your ggplot code for question 3 here
```

### Question 4 - Research Question

Oh no! Your try-hard problematic Gen X boss that wants to be a Gen Z just texted you:

> "OMG - I don't know how Trump lost in 2020 when he won so many more counties than Biden did!! I told my boss I would blow her mind with a pie chart I made about this.  Can you put one together for my meeting with her in 30 minutes?  And we need a chart that EATS: title, value labels and appropriate colors.  Thanks king!"

What do you send your boss?

**BUTTERUP YOUR BOSS BONUS (not required):** Draft an email to your boss giving them the chart, but with a warning about why this isn't the most honest way to show the 2020 results. (up to 2 points)

```{r}
# insert and run your ggplot code for question 4 here
```
