---
title: "Problem Set 3 - PPOL 6802 Spring 2024"
author: "INSERT NAME HERE"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Set #3

Here is where you will provide the answers to problem set 3. I will provide you with the basic "buckets" to put everything in to make the R Markdown creation process as straightforward as possible. But remember that if you run into any issues, ChatGPT can be extremely helpful!

### Loading the Data

To complete this problem set, you will need to download the necessary data from Github. You can do that by running the code below:

```{r message=FALSE}
library(tidyverse)

d1 <- read_csv("https://raw.githubusercontent.com/alexlundry/ppol6802_spring24/main/problem_sets/datasets/pset1_county_data.csv")

d2 <- read_csv("https://raw.githubusercontent.com/cphalpert/census-regions/master/us%20census%20bureau%20regions%20and%20divisions.csv")

d3 <- left_join(d1, d2, by = "State")
```

### Question 1 - Bar Chart with Summary Data

Using ggplot2, make a bar chart that shows the average level of education (pct_bachelors) for counties located in each census Division.  Ensure that bars are properly sorted in order and appropriately labelled.  You must also have meaningful and accurate title and axis labels.  

**Bonus (not required)**: Color the bars by their Census Region. (+1 point)

```{r echo=TRUE}
# insert and run your ggplot code in this chunk

```

### Question 2 - Grouped boxplot

Using ggplot2, create a horizontal boxplot of `pct_soc_sec` that is grouped by census Region.  Ensure that the boxplots are meaningfully sorted.  You must also have a meaningful and accurate title and axis labels. Color the boxplots by Region, but make sure there is NOT a legend displayed.  

```{r echo=TRUE}
# insert and run your ggplot code in this chunk

```

### Question 3 - Plot Reproduction

Using ggplot2, reproduce the plot below. To help you do this, here are a few key notes:

-   I have filtered out rows where `winner` is NA.
-   I am visualizing `med_income` and coloring by `winner`
-   These are density plots with transparency = 0.4
-   The plots are faceted by two variables, `Region` as the columns, and `winner` as the rows.
-   The scales of the X axis are fixed, and the scale is logged.
-   The X axis labels are formatted for displaying dollars, and the results are in thousands with the "K" suffix.
-   There is a title and axis labels.
-   There is no legend.

```{r echo=TRUE}
# insert and run your ggplot code in this chunk

```

### Question 4 - Research Question

Oh no! Your try-hard problematic Gen X boss that wants to be a Gen Z just texted you:

> "Angela and Kevin over in Accounting asked me if we could whip up a chart showing how each census division contributes to the overall population of a census region.  He said it was something about an IRS audit?? ROFL!  I told him I'd put my best people on it, but he said something about "plausible deniability"?!? "¯_(:)_/¯" Get that to me ASAP - I've got an Orange Theory class in an hour."  

What do you send your boss? (don't overthink this - the chart can be very basic)

**BUTTERUP YOUR BOSS BONUS (not required):** Offer an alternative way to visualize the data they are asking for and make a recommendation for the most effective one to use.  (up to +3 points)

```{r echo=TRUE}
# insert and run your ggplot code in this chunk

```

